What We Eval in the Shadows

### 2021-04-14 paper/main.tex
intro/abstract

\documentclass[screen,acmsmall]{acmart}
\settopmatter{printfolios=true,printccs=true,printacmref=true}

@@

\begin{abstract}
  \noindent Most dynamic languages allow users to turn text into code
  using various functions, often named \eval, with language-dependent
  semantics. The widespread use of these reflective functions hinders
  static analysis and prevents compilers from performing
  optimizations. This paper aims to provide a better sense of why
  programmers use \eval. Understanding why \eval is used in practice
  is key to finding ways to mitigate its negative impact. We have
  reasons to believe that reflective feature usage is language and
  application domain specific; we focus on data science code written
  in R, and compare our results to previous work that analyzed web
  programming in JavaScript. Dynamic analysis of a corpus of
  \CranPackages libraries and \KaggleCode end-user code confirms that
  \eval is indeed in widespread use; R's \eval is more dangerous in
  some ways, and safer in others, than what was previously reported
  for JavaScript.
  %% TODO ::: Add a sentence about summarizing why users resort to
  %% eval -- Jan
\end{abstract}

@@

\vspace{2mm}\noindent\emph{The Here and Now.} In this study, we investigate the
usage of \eval in programs written in the R programming language. R is language
designed by statisticians for applications in data science~\cite{r,R96}. What
makes looking at R after JavaScript interesting is that, while both languages
are dynamic, they are quite different. While one can program in an
object-oriented style in R, like in JavaScript, R is primarily a lazy, untyped,
functional language. JavaScript was designed to run untrusted code in browser,
while R is used for statistical computing on desktops. JavaScript is a general
purpose language used by a wide community of programmers; while R is used for
scientific computing by data scientists and domain experts with, often, limited
programming experience. One can distinguish between library implementers,
developers with some programming experience and a working knowledge of R, and
end-users, who are typically not expert programmers and often have only a
cursory knowledge of the language.\footnote{Consider that R evaluates function
argument lazily, just like Haskell. We informally surveyed end-users, including
computer scientists, and did not find a single user aware of this fact. Library
developers, on the other hand, know about laziness and program defensively
around it.} Our goal is thus to highlight the differences in usage between
JavaScript and R, and try to explain those differences in terms of language
features, application domain and programmer experience. Hopefully some of our
observations will generalize to other languages.

\vspace{2mm}\noindent\emph{The What and How.} One significant benefit of
choosing R is that every package in the CRAN repository is curated and comes
with examples of typical usage. This gives us a large code base that we can
analyze dynamically. To observe \eval we built a two-level monitoring
infrastructure:\footnote{Our infrastructure is open source and publicly
available, our anonymized code is at {\tt http://url.com}, a more complete
artifact will be submitted to the artifact evaluation committee. } we can
monitor R programs by instrumentation --- this gives us access to many
user-visible properties of R programs --- but we can also monitor the
inner-workings of the R interpreter --- this allows us to capture details not
exposed at the source level. Dynamic analysis is limited, it can only observe
behaviors triggered by the particular inputs passed to a program. Luckily, CRAN
libraries come with many tests and use-cases. The choice of corpus is crucial.
Our corpus has been constructed to reflect the levels of sophistication of the R
community. We distinguish between \emph{CRAN packages} (\CranPackages curated
packages that pass stringent quality checks and are equipped with tests and
sample data) and \emph{Kaggle scripts} (\KaggleUnique end-user written programs
that performs a particular data analysis task). It is reasonable to expect that
\eval usage differ between these datasets: the libraries represent a lively
ecosystem with new libraries added each day, while end user code is often thrown
together, run once, and never revisited.

>>>>

\documentclass[review,nonacm,screen,acmsmall,anonymous=true]{acmart}
\settopmatter{printfolios=true,printccs=true,printacmref=true}

@@

\begin{abstract}
  \noindent Most dynamic languages allow users to turn text into code using
  various functions, often named \eval, with language-dependent semantics. The
  widespread use of these reflective functions hinders static analysis and
  prevents compilers from performing optimizations. This paper aims to provide a
  better sense of why programmers use \eval. Understanding why \eval is used in
  practice is key to finding ways to mitigate its negative impact. We have
  reasons to believe that reflective feature usage is language and application
  domain specific; we focus on data science code written in R, and compare our
  results to previous work that analyzed web programming in JavaScript. This
  paper studied \CranRunnableScripts scripts extracted from \CranPackages R
  packages, for a total of \packageAllcalls calls to \eval. We find that \eval
  is indeed in widespread use; R's \eval is more pervasive and arguably
  dangerous than what was previously reported for JavaScript.
\end{abstract}

@@

\vspace{2mm}\noindent\emph{The Here and Now.} In this study, we investigate the
usage of \eval in programs written in the R programming language. R is language
designed by statisticians for applications in data science~\cite{r,R96}. What
makes looking at R after JavaScript interesting is that, while both languages
are dynamic, they are quite different. While one can program in an
object-oriented style in R, like in JavaScript, R is primarily a lazy, untyped,
functional language. JavaScript was designed to run untrusted code in browser,
while R is used for statistical computing on desktops. JavaScript is a general
purpose language used by a wide community of programmers; while R is used for
scientific computing by data scientists and domain experts with, often, limited
programming experience. One can distinguish between library implementers,
developers with some programming experience and a working knowledge of R, and
end-users, who are typically not expert programmers and often have only a
cursory knowledge of the language.\footnote{Consider that R is lazy like
Haskell. We informally surveyed end-users and did not find a single user aware
of this fact. Library developers, on the other hand, know and program
defensively around laziness.} Our goal is thus to highlight the differences in
usage between JavaScript and R, and try to explain those differences in terms of
language features, application domain and programmer experience. Hopefully some
of our observations will generalize to other languages.

\vspace{2mm}\noindent\emph{The What and How.} One significant benefit of
choosing R is that every package in the CRAN repository is curated and comes
with examples of typical usage. This gives us a large code base that we can
analyze dynamically. To observe \eval we built a two-level monitoring
infrastructure:\footnote{Our infrastructure is open source and publicly
available and  will be submitted to the artifact evaluation committee. } we can
monitor R programs by instrumentation --- this gives us access to many
user-visible properties of R programs --- but we can also monitor the
inner-workings of the R interpreter --- this allows us to capture details not
exposed at the source level. Dynamic analysis is limited, it can only observe
behaviors triggered by the particular inputs passed to a program. Luckily, CRAN
libraries come with many tests and use-cases. The choice of corpus is crucial.
Our corpus has been constructed to reflect the levels of sophistication of the R
community. We distinguish between \emph{CRAN packages} (\CranPackages curated
packages that pass stringent quality checks and are equipped with tests and
sample data) and \emph{Kaggle scripts} (\KaggleUnique end-user written programs
that performs a particular data analysis task). It is reasonable to expect that
\eval usage differ between these datasets: the libraries represent a lively
ecosystem with new libraries added each day, while end user code is often thrown
together, run once, and never revisited.
